---
title: "PA1_template.Rmd"
author: "AG"
date: "Sunday, October 11, 2015"
output: html_document
---
#Overview#


*It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a [Fitbit](http://www.fitbit.com/), [Nike] (http://www.nike.com/)[Fuelband](http://www.fuelband.com/), or [Jawbone](http://www.jawbone.com/) Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.*

This document tries to illustrate the average steps taken in a day , daily pattern and also tries to explore if there differences during weekdays and weekends. This is based of the the ["Activity" dataset](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip), provided as a part of Assignment 1 of Reproducibel Research course at JHU.

## Let us read the dataset now ##
I've set the working directory to the location where the activity.csv file is located. Before we read the dataset, want to load the *ggplot2* library for plotting. You can see the code below

```{r echo=TRUE}
library(ggplot2)
library(chron)
activitydata <-read.csv("./activity.csv", head=TRUE, sep=",")
```


### Let us look at the activitydata ###


```{r, echo=TRUE}
head(activitydata)
sapply(activitydata, class)
```


## Question 1: What is mean total number of steps taken per day? ##


Data is collected at specific intervals, everyday for 2 months. We need to sum all the steps taken per day, before we get the mean and corresponding median. 


```{r, echo=TRUE}
stepsbyday<-tapply(activitydata$steps, activitydata$date,sum)
```


Now let us plot the steps by day

``` {r, echo = TRUE}
g<-ggplot(as.data.frame(stepsbyday), aes(stepsbyday))
g+geom_histogram(binwidth = 1000, col="green") + labs( x="Steps taken by Day", y="Number of Days", title = "Steps Per Day")
```

Also, the average number of steps across days is found to be



```{r, echo=TRUE}
mean(stepsbyday, na.rm=TRUE)
```


And the median of number of steps is


``` {r, echo=TRUE}
median(stepsbyday, na.rm=TRUE)
```

## Question 2: What is the average daily activity pattern? ##

So, we need to see how people walk around, on an average, during the day.. First, let us clean up the data a bit

``` {r, echo=TRUE}
cleandata<-subset(activitydata, !is.na(activitydata$steps))
avgdailypattern<-tapply(cleandata$steps, cleandata$interval, mean)
```

Let us plot the time series data now

```{r, echo=TRUE}
plot(avgdailypattern, type="l", col="green")
abline(v=which.max(avgdailypattern)[[1]], col ="red", lwd=2)
abline(h=max(avgdailypattern), col="red", lwd=2)
```
Also we observe that the maximum number of steps are on the below intervals
```{r, echo=TRUE}
which.max(avgdailypattern)[[1]]
```

## Question 3 Imputing Missing Values ##

As the first part of the question, let us look at the number of missing values in the data set

``` {r, echo=FALSE}
sapply(activitydata, function(x) sum(is.na(x)))
````

So, all we need is to compute the missing values for steps feature. Let us see how many days have NA values

``` {r, echo=TRUE}
NADays<-tapply(activitydata$steps, activitydata$date, function(x) sum(is.na(x)))
```
Let us see how many days have NA values
```{r, echo=TRUE}
length(which(NADays!=0))
```

Let us impute values into the data set, now. Will adopt a very simple technique of inserting "mean" number of steps / interval into the data set

Let us massage the data set
```{r, echo=TRUE}
data1<-subset(activitydata, !is.na(activitydata$steps))
data2<-subset(activitydata, is.na(activitydata$steps))
steppat<-tapply(data1$steps, data1$interval, mean)
data2$steps<-steppat
data3<-rbind(data1, data2)
stepsbyday1<- tapply(data3$steps, data3$date, sum)
```

``` {r, echo = TRUE}
g<-ggplot(as.data.frame(stepsbyday1), aes(stepsbyday1))
g+geom_histogram(binwidth = 1000, col="green") + labs( x="Steps taken by Day", y="Number of Days", title = "Steps Per Day")
```

Mean and Meadian values, below
```{r, echo=TRUE}
mean(stepsbyday1)
median(stepsbyday1)
```

As you can see from above there is no change from imputing values in the dataframe

## Question 4: Are there differences in activity patterns between weekdays and weekends? ##

Let us create a factor variable to indicate weekdays vs weekends

``` {r, echo=TRUE}
activitydata$date<-as.POSIXlt(activitydata$date, format="%Y-%m-%d")
activitydata$weekend<-is.weekend(activitydata$date)
table(activitydata$weekend)
```

Let us plot the weekday Vs Weekend pattern
```{r, ech=TRUE}
activitydata<-subset(activitydata, !is.na(activitydata$steps))
weekdaypat<-subset(activitydata, !activitydata$weekend)
weekendpat<-subset(activitydata, activitydata$weekend)
wkdyavg<-tapply(weekdaypat$steps, weekdaypat$interval, mean)
wkndavg<-tapply(weekendpat$steps, weekendpat$interval, mean)
```

Let us plot the patterns

``` {r, ech=TRUE}
par(mfrow=c(2,1), mar= c(4,4,4,2))
plot(wkdyavg, type="l", col="green", title="Weekdays")
plot(wkndavg, type="l", col="green", title="Weekends")











































